# Example 1: Jester5k

This example is designed to illustrate the process of building recommender systems. The dataset we will be investigating is the Jester5k data, which contains a sample of 5000 users from the anonymous ratings data from the Jester Online Joke Recommender System collected between April 1999 and May 2003. The data set contains ratings for 100 jokes on a scale from −10 to +10. 

The dataset is available in the `recommenderlab` package:

```{r warning=FALSE}
library(recommenderlab)
data(Jester5k)
```

> Reference: [Section 5.4 of the `recommenderlab` package vignette](https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf)

## Exploratory data analysis

First, let's investigate the data a bit. Revise the R code in the lecture note and answer the following questions:

* How many items (i.e. jokes) are there in this dataset? `r fitb(100)`

* On average, how many jokes do the users rate? You can report the median value in this case to get an integer value. `r fitb(72)`

* On average, how many ratings have been made to each joke? `r fitb(3632)`

`r hide("Hint")`
You can get information about the dataset by using the `cat()` function. `rowCounts` and `colCounts` can be useful to get information about items and users. 
`r unhide()`

```{r warning=FALSE, webex.hide="Solution"}
cat(JesterJokes[1]) #Joke information

Jester5k
Jester5k@data[1:5,1:5]
image(Jester5k@data[1:5,1:5])

summary(rowCounts(Jester5k)) #No. jokes rated by a user
summary(colCounts(Jester5k)) #No. users rated a joke

hist(getRatings(Jester5k), breaks=100) #getRating(): extract a vector with all non-missing ratings from a rating matrix
```

##Build recommender systems and make predictions

First, let's build a user-based collaborative filtering (UBCF) system to predict the ratings and the top-$N$ recommendations. Here, we use Pearson correlation to measure the similarity between users and select the nearest 5 users to form a neighbourhood. 
```{r}
Jester_UBCF <- Recommender(Jester5k, method="UBCF",
                           param=list(method="pearson",nn=5))

# Predict ratings only for unrated items
Jester_UBCF_rating <- predict(Jester_UBCF, Jester5k, type="ratings")
Jester_UBCF_rating@data[5:10,1:10] 

# Predict ratings for all items 
Jester_UBCF_ratMat <- predict(Jester_UBCF, Jester5k, type="ratingMatrix")
Jester_UBCF_ratMat@data[5:10,1:10] 

# Recommend top-3 items.
Jester_UBCF_topN <- predict(Jester_UBCF, Jester5k, type="topNList", n=3)
getList(Jester_UBCF_topN)[1:5]
# some users have rated all items and thus the prediction returns "character(0)"
```

Let's now move on to an item-based collaborative filtering (IBCF) system.
```{r eval=FALSE}
Jester_IBCF <- Recommender(Jester5k, method="IBCF", param=list(k=20)) #k: no. similar items
```

You can revise the previous codes on the UBCF system to predict ratings and top-$N$ items with the new IBCF system. 

Finally, we can perform UV-decomposition on the rating matrix. Strictly speaking, the package implements the idea of matrix factorization using singular value decomposition (SVD), which is numerically more robust and efficient. The key parameter for matrix decomposition (either UV or SVD) is the latent dimension, which is set as 3 in the following code. 
```{r}
Jester_UV <- Recommender(Jester5k,method="SVDF", param=list(k=3)) #k: latent dimension
dim(Jester_UV@model$svd$U)
dim(Jester_UV@model$svd$V)
```


## Evaluate recommender systems using built-in functions

### Evaluate predicted ratings

We evaluate the three recommendation systems using 10-fold cross-validation. For the test set, 5 items will be given to the recommender algorithm to make the prediction and the rest of items will be held out for computing the error. Below is an example on evaluating UBCF. 
```{r}
set.seed(1)
Jester_eval <- evaluationScheme(Jester5k, method="cross-validation", train=10, given=5)
Jester_eval

UV_results <- evaluate(Jester_eval, method="UBCF", type="ratings",
                       param=list(method="pearson", nn=20))
getResults(UV_results)
avg(UV_results)
```

When evaluating multiple recommender algorithms, we can create a list of algorithms and evaluate them all together. The following R code will take a while to complete. 

```{r} 
algs <- list(
  "UBCF" = list(name="UBCF", param=list(method="pearson",nn=20)),
  "IBCF" = list(name="IBCF", param=list(k=5)),
  "UV"   = list(name="SVDF", param=list(k=5))
)
Jester_results <- evaluate(Jester_eval, algs, type="ratings")

# getResults(Jester_results$UBCF)
# getResults(Jester_results$IBCF)
# getResults(Jester_results$UV)
avg(Jester_results)
plot(Jester_results, legend="topright")
```

### Evaluate top-N recommendations

Given our data is a real rating matrix, we need to convert it into a binary matrix in order to evaluate if the top-$N$ recommended items will be liked by the user. This conversion is achieved by using the argument `goodRating`, which is a threshold on the actual ratings; values at or above the threshold is considered to be liked by the user and vice versa. In other words, an item in the topNList is considered a true positive if it has a rating of `goodRating` or better in the observed data.

```{r}
Jester_eval2 <- evaluationScheme(Jester5k, method="cross", k=5, given=-5,
                                 goodRating=0) 
Jester_eval2
Jester_results2 <- evaluate(Jester_eval2, algs, type="topNList",
                            n=c(1,3,5,10))
# n: the number of items in top-N list

avg(Jester_results2)
plot(Jester_results2, annotate=1, legend="topleft")
```

The above example covers collaborative filtering recommender systems. If you are interested in content-based recommender systems, the following example may be studied: 
https://michael.hahsler.net/other_courses/ICMA_Recommendation_Tools/code/content-based.html#calculate-content-based-item-similarity

