---
title: "STATS5099 Data Mining and Machine Learning"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

```{r setup, include = FALSE}
library(webexercises)
library(tinytex)
library(dplyr)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)

library(reticulate)
```


```{r include=FALSE, echo=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Welcome to DMML Lab 10

In week 10, we studied convolutional neural networks (CNNs) for image classification. In today's lab, we will revise the CNN architecture and apply it to MNIST, a database of handwritten digits. 

## Convolutional neural networks

In Lecture Note 10, we have constructed a CNN which consists of two convolutional layer, each followed by a rectified linear (ReLU) activation function and a max-pooling layer. After flatting patches into vectors, two fully-connected feedforward neural networks are applied. The code for such a CNN is reproduced below.

```{python}
import torch.nn as nn
import torch.nn.functional as F

class Net(nn.Module):
    def __init__(self):
        super().__init__()
        self.conv1 = nn.Conv2d(3, 6, 5)
        # 1st arg: in_channels = 3 (for RGB channels),
        # 2nd arg: number of filters = 6
        # 3rd arg: kernel size = 5
        self.pool = nn.MaxPool2d(2, 2)
        # 1st arg: kernel size = 2
        # 2nd arg: stride = 2
        self.conv2 = nn.Conv2d(6, 16, 5)
        self.fc1 = nn.Linear(16 * 5 * 5, 120)
        # 1st arg: input dimensionality; 16 channels * 5x5 patches
        # 2nd arg: output dimensionality
        self.fc2 = nn.Linear(120, 84)
        self.fc3 = nn.Linear(84, 10)

    def forward(self, x):
        x = self.pool(F.relu(self.conv1(x)))
        x = self.pool(F.relu(self.conv2(x)))
        x = torch.flatten(x, 1) # flatten all dimensions except batch
        x = F.relu(self.fc1(x))
        x = F.relu(self.fc2(x))
        x = self.fc3(x)
        return x

net = Net()
```

You may visit the PyTorch website to learn more details on [2D convolution](https://pytorch.org/docs/stable/generated/torch.nn.Conv2d.html) or [2D max-pooling](https://pytorch.org/docs/stable/generated/torch.nn.MaxPool2d.html). 